---
title: "Introduction to ESG"
author: "Zach Eisenstein"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
  # toc_depth: 2
vignette: >
  %\VignetteIndexEntry{intro-to-ESG}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

To Do:

* funnel graph function

```{r, include = FALSE}

library(tidyverse)
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.width = 8, 
                      fig.height = 4.5,
                      fig.align = 'center',
                      out.width='95%', 
                      dpi = 200)

# library(ESG)
devtools::load_all()
```

# Introduction

This package provides functions for generating economic scenario sets for insurance Economic Capital Models.

The two main objectives are calibration and simulation of financial variables including: 

* Inflation
* Interest Rates (real & nominal)
* Equity Returns

This package relies heavily on the `tidyverse`.

Some datasets are available with the package. Those can be accessed using the `data` command:

```{r}
data(package = "ESG")$results[, "Item"]
```


## Inflation

### Calibration

Here we will use historical inflation data to calibrate a short rate model. 

```{r}
# Load dataset
data(inflation)

head(inflation)

```

The tibble includes monthly CPI and annual inflation back to 1948. For more info see `help(inflation)`

```{r}
inflation %>%
  ggplot(aes(date, inf)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Historical Inflation",
       y = "Inflation") +
  theme_classic()
```

Now that we have our data, we can calibrate a model to project future inflation.
Here I will use the full dataset to fit a Vasicek model, which is mean reverting one-factor short rate model. For more see [here](https://en.wikipedia.org/wiki/Vasicek_model)

```{r}
inf_parm <- CalVasicek1f(inflation$inf)

inf_parm
```
The result is a list with the following parameters:

* **r0** - the starting value (last value of the data provided)
* **m** - mean reverting level
* **k** - strength of the mean reversion (annual)
* **v** - volatility (annual) 

### Simulation

Now that we have our parameters, we can simulate future inflation paths. We will do this with a call to `SimVasicek1F`

```{r}
inf_sim <- SimVasicek1F(n=10, inf_parm)

inf_sim

inf_sim %>%
  bind_rows(tibble(trial = 1:10, time = 0, value = inf_parm$r0)) %>%
  ggplot(aes(time, value, group = trial)) +
  geom_line(alpha = 1/2) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Future Projected Inflation",
       y = "Inflation") +
  theme_classic()

```
Data is simulated in monthly increments with a default time of 1 year. 
If we're interested in projections further out we can adjust the time parameter `t`

```{r}
inf_sim <- SimVasicek1F(n=10, inf_parm, t = 60)

inf_sim

inf_sim %>%
  bind_rows(tibble(trial = 1:10, time = 0, r = inf_parm$r0)) %>%
  ggplot(aes(time, value, group = trial)) +
  geom_line(alpha = 1/2) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Future Projected Inflation",
       y = "Inflation") +
  theme_classic()

```

Now let's simulate 1000 paths and view output.
Convenience function `quibble` provides some summary statistics.

```{r}
inf_sim <- SimVasicek1F(n=1e3, inf_parm)

inf_sim %>%
  mutate(time = round(time, 3)) %>%
  group_by(time) %>%
  summarise(quibble(value)) %>%
  pivot_wider(names_from = time)
```

We can also visualize the range of distributions using a funnel graph `gg_funnel`.
This shows historical values alongside median and various quantiles of the future forecast.

```{r}

gg_funnel(inf_sim, tail(inflation$inf, 61), inf_parm$r0, variable_name = "Inflation", time_offset = 2022)

```


## Equities

